M.Sc. Proposal

### 2010-09-26 notes/proposal.tex
Review fixes.

\title{M.Sc. Proposal}
\author{Roman Zeyde \\ \\  {\small Advisor: Prof. Irad Yavneh} }
\begin{document}
\maketitle
\section{Introduction}
Electrokinetic theory describes the dynamics of charged particles
in ionic fluids. When a particle acquires surface charge, a layer
of ions of opposite charge is attracted to the surface via
electric forces, creating a double-layer structure around the
particle (see Figure \ref{fig:EDL}). This structure, called
``Debye layer'', electrically screens the surface charge, thus
creating a potential difference between the particle and the outer
layer of the fluid bulk.

\begin{figure}[htbp]
\begin{framed}
    \begin{center}
        \includegraphics[width=0.25\textwidth]
            {ElectricDoubleLayer.eps}
        \caption{Schematic structure of the double layer:
        (1) particle surface, (2) Debye layer and (3) fluid bulk.
        If the particle surface has positive (blue) surface charge,
        it attracts negative (red) ions from the fluid making the
        Debye layer negatively charged (as opposed to the rest of
        the fluid bulk, which is electrically neutral).
        Zeta potential is defined as the voltage drop on (2).}
        \label{fig:EDL}
    \end{center}
\end{framed}
\end{figure}
In cases where the layer width is much smaller than the particle's
size, an analytical asymptotic solution for the Debye layer
dynamics can be found \cite{Yariv10a}.

Electrokinetics is described by the following physical phenomena (in dimensionless formulation):
\begin{enumerate}
\item Electrostatic potential---the fluid bulk contains no free
charge:
\begin{equation} \label{eq:Laplace}
    \bnabla \cdot (C \bnabla \varPhi) = 0 \, .
\end{equation}
\item Fluid dynamics---Stokes flow with an electrostatic force
component:
\begin{equation} \label{eq:Stokes}
\left\{ \begin{array}{l}
\bLaplacian \bV - \bnabla P + \Laplacian \varPhi \bnabla \varPhi = 0 \, , \\
\bnabla \cdot \bV = 0 \, . \end{array} \right.
\end{equation}
\item Ion diffusion and advection (via the fluid's velocity field):
\begin{equation} \label{eq:Nernst}
\Laplacian C - \alpha \bV \cdot \bnabla C = 0 \, .
\end{equation}
\end{enumerate}
{\bf IY: WRITE EXPLICITLY HERE WHAT THE VARIABLES ARE---C,P,V,
ETC.}The boundary conditions are determined by the specific
problem under consideration and defined by the particle's
geometry, chemical characteristics, and the fluid dynamics.

@@

{\bf IY: ADD HERE A PARAGRAPH THAT DISCUSSES THE IMPORTANCE AND
RELEVANCE OF THE PROBLEM---YOU CAN TAKE THIS FROM THE PAPERS YOU
HAVE READ.}

\section{Research goals}
Our goal is to implement an accurate and fast iterative numerical
solver for electrokinetic problems, and then apply it to the study
of systems that have no closed-form solutions, including strong
electric fields, non-spherical particles, and asymmetric charge on
the particle surface. We expect to thus gain insight into the
chemical and physical behavior in far more general regimes than
are currently well-understood. We hope that this will also lead to
further analytical developments, based on new scaling behavior
provided by the numerical solver. We may also target interesting
phenomena that have been observed experimentally, e.g.,
electrophoretic\footnote{Electrophoresis is the physical
phenomenon of the motion of particles in an ion-containing fluid
under the influence of an electric field.} autonomous
micro-swimmers \cite{Paxton04, Howse07}.

@@

\section{Preliminary results}
The numerical solver is being implemented as a MATLAB package,
which provides the needed tools for numerical solution of
electrokinetic problems. The solver applies an iterative
relaxation technique to solve the discretized coupled system.
Beginning with some initial guess for the solution The equations
are relaxed in order with respect to their corresponding
variables, with the remaining variables fixed at their current
values (``Gauss-Seidel style''):
\begin{itemize}
  \item The diffusion equation (\ref{eq:Laplace}) is used to update $\varPhi$.
  \item The Stokes system (\ref{eq:Stokes}) is used to update $\bV$ and $P$.
  \item The Nernst-Planck equation (\ref{eq:Nernst}) is used to update $C$.
\end{itemize}

@@

\begin{thebibliography}{}
\bibitem{Yariv10a} E. Yariv,
``An asymptotic derivation of the thin-debye-layer limit for electrokinetic phenomena'',
{\em Chem. Engng Comm. 197, 3–17}, 2010.
\bibitem{Yariv10b} E. Yariv,
``Migration of ion-exchange particles driven by a uniform electric field'',
{\em Journal of Fluid Mechanics, 655 105-121}, 2010.
\bibitem{Rubin01} I. Rubinstein and B. Zaltzman,
``Electro-osmotic slip of the second kind and instability in concentration polarization at electrodialysis membranes'',
{\em Math. Mod. Meth. Appl. Sci. 11 (2), 263–300}, 2001.
\bibitem{Paxton04}
W. Paxton et. al.,
``Catalytic Nanomotors: Autonomous Movement of Striped Nanorods'',
{\em Journal of the American Chemical Society 126 (41), 13424-13431}, 2004.
\bibitem{Howse07}
J. R. Howse et. al.,
``Self-Motile Colloidal Particles: From Directed Propulsion to Random Walk'',
{\em Phys. Rev. Lett., 99, 048102}, 2007.
\bibitem{Vanka86}
S. P. Vanka,
``Block-implicit multigrid solution of Navier-Stokes equations in primitive variables'',
{\em J. Copmput. Phys., 65, 138-158}, 1986.
\bibitem{Sidi90}
A. Sidi, ``Efficient implementation of minimal polynomial and reduced rank extrapolation methods'',
{\em Journal of Computational and Applied Mathematics, Vol. 36, pp. 305-337}, 1991.
\bibitem{Brandt77} Achi Brandt,
``Multi-Level Adaptive Solutions to Boundary-Value Problems'',
{\em Mathematics of Computation, 31: 333–90}, 1977.
\bibitem{Yavneh06} Irad Yavneh.
``Why Multigrid Methods are so Efficient'',
{\em Computing in Science and Engineering, 8(6):12-22}, 2006
\end{thebibliography}

>>>>

\title{M.Sc. Proposal}
\author{Roman Zeyde \\ \\  {\small Advisor: Prof. Irad Yavneh} }
\begin{document}
\maketitle
\section{Introduction}
Electrokinetic theory describes the dynamics of charged particles
in ionic fluids. When a particle acquires surface charge, a layer
of ions of opposite charge is attracted to the surface via
electric forces, creating a double-layer structure around the
particle (see Figure \ref{fig:EDL}). This structure, called
``Debye layer'', electrically screens the surface charge, thus
creating a potential difference between the particle and the outer
layer of the fluid bulk.
\begin{figure}[htbp]
\begin{framed}
    \begin{center}
        \includegraphics[width=0.25\textwidth]
            {ElectricDoubleLayer.eps}
        \caption{Schematic structure of the double layer:
        (1) particle surface, (2) Debye layer and (3) fluid bulk.
        If the particle surface has positive (blue) surface charge,
        it attracts negative (red) ions from the fluid making the
        Debye layer negatively charged (as opposed to the rest of
        the fluid bulk, which is electrically neutral).
        Zeta potential is defined as the voltage drop on (2).}
        \label{fig:EDL}
    \end{center}
\end{framed}
\end{figure}
In cases where the layer width is much smaller than the particle's
size, an analytical asymptotic solution for the Debye layer
dynamics can be found \cite{Yariv10a}.

Electrokinetics is described by the following physical phenomena (in dimensionless formulation):
\begin{enumerate}
\item Electrostatics---the fluid bulk contains no free
charge:
\begin{equation} \label{eq:Laplace}
    \bnabla \cdot (C \bnabla \varPhi) = 0 \, .
\end{equation}
\item Fluid dynamics---Stokes flow with an electrostatic force
component:
\begin{equation} \label{eq:Stokes}
\left\{ \begin{array}{l}
\bLaplacian \bV - \bnabla P + \Laplacian \varPhi \bnabla \varPhi = 0 \, , \\
\bnabla \cdot \bV = 0 \, . \end{array} \right.
\end{equation}
\item Ion diffusion and advection (via the fluid's velocity field):
\begin{equation} \label{eq:Nernst}
\Laplacian C - \alpha \bV \cdot \bnabla C = 0 \, .
\end{equation}
\end{enumerate}
The variables of the electrokinetic problem are the electrostatic 
potential $\varPhi$, fluid's velocity $\bV$ and its pressure $P$ 
and ionic concentration $C$.

The boundary conditions are determined by the specific
problem under consideration and defined by the particle's
geometry, chemical characteristics, and the fluid dynamics.

@@

Recently, it has been conceived that such type of dynamics 
can be very useful in transporting and manipulating micro- 
and nanoscale objects for many nanotechnology applications, 
such as the self-assembly of superstructures, roving sensors, 
drug-delivery systems and useful nanomachinery \cite{Pumera10}. 
However, since the analytic model for these electrokinetic 
phenomena has no closed-form analytic solution, a numerical
solver has to be developed.

\section{Research goals}
Our goal is to implement an accurate and fast iterative numerical
solver for electrokinetic problems, and then apply it to the study
of systems that have no closed-form solutions, including strong
electric fields, non-spherical particles, and asymmetric charge 
distribution on the particle surface. We expect to thus gain insight 
into the chemical and physical behavior in far more general regimes 
than are currently well-understood. We hope that this will also lead 
to further analytical developments, based on new scaling behavior
provided by the numerical solver. We may also target interesting
phenomena that have been observed experimentally, e.g.,
electrophoretic\footnote{Electrophoresis is the physical
phenomenon of the motion of particles in an ion-containing fluid
under the influence of an electric field.} autonomous
micro-swimmers \cite{Paxton04, Howse07}.

@@

\section{Preliminary results}
The numerical solver is being implemented as a MATLAB package,
which provides the needed tools for numerical solution of
electrokinetic problems. The solver applies an iterative
relaxation technique to solve the discretized coupled system.
Beginning with some initial guess for the solution, the equations
are relaxed in order with respect to their corresponding
variables, with the remaining variables fixed at their current
values (``Gauss-Seidel style''):
\begin{itemize}
  \item The diffusion equation (\ref{eq:Laplace}) is used to update $\varPhi$.
  \item The Stokes system (\ref{eq:Stokes}) is used to update $\bV$ and $P$.
  \item The Nernst-Planck equation (\ref{eq:Nernst}) is used to update $C$.
\end{itemize}

@@

\begin{thebibliography}{}
\bibitem{Yariv10a} E. Yariv,
``An asymptotic derivation of the thin-debye-layer limit for electrokinetic phenomena'',
{\em Chem. Engng Comm. 197, 3–17}, 2010.
\bibitem{Yariv10b} E. Yariv,
``Migration of ion-exchange particles driven by a uniform electric field'',
{\em Journal of Fluid Mechanics, 655 105-121}, 2010.
\bibitem{Pumera10} M. Pumera,
``Electrochemically powered self-propelled electrophoretic nanosubmarines''
{\em Nanoscale, 2, 1643-1649}, 2010.
\bibitem{Rubin01} I. Rubinstein and B. Zaltzman,
``Electro-osmotic slip of the second kind and instability in concentration polarization at electrodialysis membranes'',
{\em Math. Mod. Meth. Appl. Sci. 11 (2), 263–300}, 2001.
\bibitem{Paxton04}
W. Paxton et. al.,
``Catalytic Nanomotors: Autonomous Movement of Striped Nanorods'',
{\em Journal of the American Chemical Society 126 (41), 13424-13431}, 2004.
\bibitem{Howse07}
J. R. Howse et. al.,
``Self-Motile Colloidal Particles: From Directed Propulsion to Random Walk'',
{\em Phys. Rev. Lett., 99, 048102}, 2007.
\bibitem{Vanka86}
S. P. Vanka,
``Block-implicit multigrid solution of Navier-Stokes equations in primitive variables'',
{\em J. Copmput. Phys., 65, 138-158}, 1986.
\bibitem{Sidi90}
A. Sidi, ``Efficient implementation of minimal polynomial and reduced rank extrapolation methods'',
{\em Journal of Computational and Applied Mathematics, Vol. 36, pp. 305-337}, 1991.
\bibitem{Brandt77} Achi Brandt,
``Multi-Level Adaptive Solutions to Boundary-Value Problems'',
{\em Mathematics of Computation, 31: 333–90}, 1977.
\bibitem{Yavneh06} Irad Yavneh.
``Why Multigrid Methods are so Efficient'',
{\em Computing in Science and Engineering, 8(6):12-22}, 2006
\end{thebibliography}
